Self Referencing Sequences

### 2014-11-06 draft.tex
Edits to Equivalence

We would first like to make a formal definition of equivalence. 
\begin{definition}
We say that two sequence $\{a_i\}$ and $\{b_i\}$ are equivalent if there exists some integers $n, k \geq 0$ such that 
\begin{align*}
a_{n+i} = b_{k+i}
\end{align*}
for all positive integers $i$. 
\end{definition}

As an example let us consider sequences generated by $C = (123)$. The sequence which starts with a 2 is 
\begin{align*}
2,2,3,3,1,1,1,2,2,2,3,1,2,3,3,1,1,2,2,3,3,3,\dots \ ,
\end{align*}
and the sequence starting with 3 is 
\begin{align*}
3,3,3,1,1,1,2,2,2,3,1,2,3,3,1,1,2,2,3,3,3,\dots \ .
\end{align*}
If we delete the first two numbers of the sequence starting with 2, and the first number of the sequence starting with 3, we see 
\begin{align*}
2,2,&|3,3,1,1,1,2,2,2,3,1,2,3,3,1,1,2,2,3,3,3,\dots\\
3,&|3,3,1,1,1,2,2,2,3,1,2,3,3,1,1,2,2,3,3,3,\dots \ ,
\end{align*}
that the resultant sequences appear to be identical and thus equivalent. At an intuitive level it seems obvious that these two truncated sequences are identical as they both begin with the same two numbers and the numbers which were removed do not refer to blocks to the right of these numbers.

In order to formalize and generalize this idea we introduce the notion of a read distance.  
\begin{definition} Consider a sequence $\{a_i\}$. We define the read distance for an index $i$ as the number of terms in between it $a_i$ and the beginning of the $i$th block of numbers.
\end{definition}
The read distance can be any nonnegative integer. For example if we consider the sequence $\{a_i\}$ generated by $C = (1 2 3)$ beginning with a 1  
\begin{align*}
1,2,2,3,3,1,1,1,2,2,2,3,1,2,3,3,1,1,2,2,3,3,3,\dots
\end{align*}
the read distance for $1$ is always $0$ since in a self-referencing sequence, the first number must refer to the first block. The read distance for $2$ is $0$ as the second block of numbers starts at the second number. We can continue to calculate the read distances for this sequence and see that they form a monotonically increasing sequence $0, 0 ,1,2,4,6,6,6,6,7,8,9,\ldots$.

This monotonicity is true in general, and we can in fact make a stronger statement on the value of the read distance at a given index.
\begin{lemma} 
\label{readdistformula}
Let $\{a_i\}$ be any self referencing sequence generated by some cycle of positive integers. Then the read distance for $i>1$, is 
\begin{equation*}
\sum_{k=1}^{i-1}( a_k - 1)
\end{equation*} 
and the read distance for $1$ is always 0.
\end{lemma}
\begin{proof}
We will prove this using induction. We know that for $i=1$ the read distance is $0$, as the first block always starts at the first number. When $i=2$ the read distance is $a_1-1$. The first block will be of length $a_1$ since the sequence is self referencing, therefore the second block will begin at position $a_1$. The number of elements between position $a_1$ and $2$ is exactly $a_1 -1$.

Now if we assume that our result holds for $i=l$ we would like to show that it holds for $i=l+1$. We know that the distance between the start of the $l$th block and position $l$ is 
\begin{equation*}
\sum_{k=1}^{l-1}( a_k - 1).
\end{equation*} 
This $l$th block has length $a_l$ and so the $l+1$th block will begin $a_l$ positions to the right of the beginning of the $l$th block. Also note that $a_{l+1}$ is one position to the right of $a_l$. Therefore the beginning of the $(l+1)$th block is 
\begin{equation*}
a_l-1 + \sum_{k=1}^{l-1} (a_k-1) = \sum_{k=1}^{l} (a_k-1)
\end{equation*} 
positions to the right of $a_{l+1}$. But this is exactly the reading distance for $a_{l+1}$ and so we have our result.
\end{proof}
Note that because $a_i$ is generated by positive integers we always have $a_k-1\geq 0$ and so the read distance is monotonically increasing.

@@

Now that we have a nice expression for the value of the read distance at a particular index we can show our first major result which allows us to check only finitely many indices in order to establish that two sequences are equivalent. 
\begin{theorem} 
\label{equivsuff}
Let $\{a_i\}, \{b_i\}$ be two self-referencing sequences generated by the same cycle.If there are integers $m,l$ with the read distance for $a_m$ and $b_l$ are both $d$ and 
\begin{equation*}
a_{m+k} = b_{l+k}
\end{equation*}
where $0 \leq k \leq d-1$ then $\{a_i\}$ and $\{b_i\}$ are equivalent. 
\end{theorem}

\begin{proof}
We will prove this by  induction. In particular we will show for all $k \geq 0$ that $a_{m+k} = b_{l+k}$ and the read distance for $a_{m+k}$ equals the read distance for $b_{l+k}$. By our assumptions we already have this for $0 \leq k \leq d-1$ and so we begin with $k=d$.

Since a new block begins at $a_{m+d}$ and $b_{l+d}$ and $a_{m+d-1} = b_{l+d-1}$ we have $a_{m+d} = b_{l+d}$. Furthermore, by Proposition ~\ref{readdistformula} we have that the read distance for $a_{m+d}$ equals the read distance for $b_{l+d}$.

Now suppose that  for $k \leq j$ we have $a_{m+k} = b_{l+k}$ and the read distance for $a_{m+k}$ equals the read distance for $b_{l+k}$. We would like to show that this holds for $k = j+1$. By Proposition ~\ref{readdistformula} we actually already know that the read distance for the two is the same, as it is the same for the preceding terms, which are equal.

If $a_{m+k}$ is the last term in a block then $b_{m+k}$ must also be the last term in its block, because the length of blocks is determined by earlier terms which we know to be equal and to have the same read distance. Since $a_{m+k} = b_{l+k}$  in this case we have $a_{m+k+1} = b_{l+k+1}$.

Otherwise $a_{m+k}$ and $a_{m+k+1}$ are in the same block, and so we only need to show that $b_{l+k}$ and $b_{l+k+1}$ are in the same block as well. But as we argued above the length of blocks is determined by terms earlier in the sequence, which we know to be equal and have the same read distance. Therefore we have $a_{m+k+1} = b_{l+k+1}$ for this case as well. This completes our induction and so $\{a_i\}$ and $\{b_i\}$ are equivalent. 
\end{proof}

This result also allows us to make a general statement for equivalence for sequences that begin with a $1$.
\begin{corollary} If $C = (1 c_1 c_2 \cdots c_n)$, then $s(C, 1)$ and $s(C, c_1)$ are equivalent.
\end{corollary}
\begin{proof}
Let $\{a_i\} = s(C, 1)$ and $\{b_i\} = s(C, c_1)$. The read distance of sequence $\{a_i\}$ for $i=2$ is 0, and the read distance of sequence $\{b_i\}$ for $i = 1$ is 0. In addition, $a_2 = b_1 = c_1$, so by Proposition ~\ref{equivsuff} the two sequences are equivalent.
\end{proof}

>>>>

We would first like to make a formal definition of equivalence. 
\begin{definition}
We say that two sequence $\{a_i\}$ and $\{b_i\}$ are equivalent if there exists some integers $n, k \geq 0$ such that 
\begin{align*}
a_{n+i} = b_{k+i},
\end{align*}
for all positive integers $i$. 
\end{definition}

As an example let us consider sequences generated by $C = (123)$. The sequence which starts with a 2 is 
\begin{align*}
2,2,3,3,1,1,1,2,2,2,3,1,2,3,3,1,1,2,2,3,3,3,\dots \ ,
\end{align*}
and the sequence that starts with a 3 is 
\begin{align*}
3,3,3,1,1,1,2,2,2,3,1,2,3,3,1,1,2,2,3,3,3,\dots \ .
\end{align*}
If we delete the first two numbers of the sequence starting with 2, and the first number of the sequence starting with 3, we see 
\begin{align*}
2,2,&|3,3,1,1,1,2,2,2,3,1,2,3,3,1,1,2,2,3,3,3,\dots\\
3,&|3,3,1,1,1,2,2,2,3,1,2,3,3,1,1,2,2,3,3,3,\dots \ ,
\end{align*}
that the resultant sequences appear to be identical and thus equivalent. At an intuitive level it seems obvious that these two truncated sequences are identical as they begin with the same two numbers and the numbers which were removed do not refer to blocks to the right of these numbers.

In order to formalize this idea we introduce the notion of a read distance.  
\begin{definition} Consider a sequence $\{a_i\}$. We define the read distance for an index $i$ as the number of terms in between it $a_i$ and the beginning of the $i$th block of numbers.
\end{definition}
The read distance can be any nonnegative integer. For example if we again consider generated by $C = (1 2 3)$ that starts with a $2$
\begin{align*}
2,2,3,3,1,1,1,2,2,2,3,1,2,3,3,1,1,2,2,3,3,3,\dots \ ,
\end{align*}
the read distance for the first number, $2$, is $0$ since in a self-referencing sequence, the first number must refer to the first block. The read distance for the second $2$ is $1$ as the second block of numbers starts at the third number. We can continue to calculate the read distances for this sequence and see that they form a monotonically increasing sequence $0 ,1,2,4,6,6,6,6,7,8,9,\ldots$.

This monotonicity is true in general, and we can in fact make a stronger statement on the value of the read distance at a given index.
\begin{lemma} 
\label{readdistformula}
Let $\{a_i\}$ be any self referencing sequence generated by some cycle of positive integers. Then the read distance for the number at index $i>1$, is 
\begin{equation*}
\sum_{k=1}^{i-1}( a_k - 1),
\end{equation*} 
and the read distance for the number at index $1$ is always 0.
\end{lemma}
\begin{proof}
We will prove this using induction. We know that for $i=1$ the read distance is $0$, as the first block always starts at the first number. When $i=2$ the read distance is $a_1-1$. The first block will be of length $a_1$ since the sequence is self referencing, therefore the second block will begin at position $a_1$. The number of elements between position $a_1$ and $2$ is exactly $a_1 -1$.

Now if we assume that our result holds for $i=l$ we would like to show that it holds for $i=l+1$. We know that the distance between the start of the $l$th block and position $l$ is 
\begin{equation*}
\sum_{k=1}^{l-1}( a_k - 1).
\end{equation*} 
This $l$th block has length $a_l$ and so the $l+1$th block will begin $a_l$ positions to the right of the beginning of the $l$th block. Also note that $a_{l+1}$ is one position to the right of $a_l$. Therefore the beginning of the $(l+1)$th block is 
\begin{equation*}
a_l-1 + \sum_{k=1}^{l-1} (a_k-1) = \sum_{k=1}^{l} (a_k-1)
\end{equation*} 
positions to the right of $a_{l+1}$. But this is exactly the reading distance for $a_{l+1}$ and so we have our result.
\end{proof}
Note that because $\{a_i\}$ is generated by positive integers we always have $a_k-1\geq 0$ and so the read distance is monotonically increasing.

@@

Now that we have a nice expression for the value of the read distance at a particular index we can show our first major result which allows us to check only finitely many indices in order to establish that two sequences are equivalent. 
\begin{theorem} 
\label{equivsuff}
Let $\{a_i\}, \{b_i\}$ be two self-referencing sequences generated by the same cycle.If there are integers $m,l$ such that the read distance for $a_m$ and $b_l$ are both $d$ and 
\begin{equation*}
a_{m+k} = b_{l+k},
\end{equation*}
for all $0 \leq k \leq d-1$ then $\{a_i\}$ and $\{b_i\}$ are equivalent. 
\end{theorem}

\begin{proof}
We will proceed by contradiction so assume otherwise. That is there exists some $j \geq d$ such that 
\begin{equation*}
a_{m+j} \neq b_{l+j}
\end{equation*}
but 
\begin{equation*}
a_{m+k} = b_{l+k}
\end{equation*} 
for all $0 \leq k < j$. 
Since the two sequences are both generated by the same cycle this implies that either $a_{m+j}$ or $b_{l+j}$ is part of the same block as the element proceeding it, and the other is not.
Without loss of generality we can assume that $a_{m+j}=a_{m+j-1}$ and $b_{l+j} \neq b_{l+j-1}$. 
The length of the block that $a_{m+j-1}$ is a part of is given by some element $a_{m+n}$, for $0 \leq n < j-1$. Because the two sequences are identical for $0 \leq k<j$ the length of the block that $b_{l+j-1}$ is a part of is given by the element $b_{l+n}$.  
We note that $a_{m+n}=b_{l+n}$ and so the two block including $a_{m+j-1}$ must be of the same length as the block including $b_{l+j-1}$. 
Since the two sequences are the same for $0 \leq k<j$ the number of elements in the block containing $a_{m+j-1}$ and in the block containing $b_{l+j-1}$ which have $k<j$ must be the same. But this means that $b_{l+j}$ is a part of the same block as the element proceeding it which is a contradiction. 
Therefore the two sequences must be equivalent.

%We will prove this by  induction. In particular we will show for all $k \geq 0$ that $a_{m+k} = b_{l+k}$ and the read distance for $a_{m+k}$ equals the read distance for $b_{l+k}$. By our assumptions we already have this for $0 \leq k \leq d-1$ and so we begin with $k=d$.

%Since a new block begins at $a_{m+d}$ and $b_{l+d}$ and $a_{m+d-1} = b_{l+d-1}$ we have $a_{m+d} = b_{l+d}$. Furthermore, by Lemma ~\ref{readdistformula} we have that the read distance for $a_{m+d}$ equals the read distance for $b_{l+d}$.

%Now suppose that  for $k \leq j$ we have $a_{m+k} = b_{l+k}$ and the read distance for $a_{m+k}$ equals the read distance for $b_{l+k}$. We would like to show that this holds for $k = j+1$. By Lemma ~\ref{readdistformula} we actually already know that the read distance for the two is the same, as it is the same for the preceding terms, which are equal.

%If $a_{m+k}$ is the last term in a block then $b_{m+k}$ must also be the last term in its block, because the length of blocks is determined by earlier terms which we know to be equal and to have the same read distance. Since $a_{m+k} = b_{l+k}$  in this case we have $a_{m+k+1} = b_{l+k+1}$.

%sOtherwise $a_{m+k}$ and $a_{m+k+1}$ are in the same block, and so we only need to show that $b_{l+k}$ and $b_{l+k+1}$ are in the same block as well. But as we argued above the length of blocks is determined by terms earlier in the sequence, which we know to be equal and have the same read distance. Therefore we have $a_{m+k+1} = b_{l+k+1}$ for this case as well. This completes our induction and so $\{a_i\}$ and $\{b_i\}$ are equivalent. 
\end{proof}

This result also allows us to make a general statement for equivalence of sequences that begin with a $1$.
\begin{corollary} If $C = (1 c_1 c_2 \cdots c_n)$, then $s(C, 1)$ and $s(C, c_1)$ are equivalent.
\end{corollary}
\begin{proof}
Let $\{a_i\} = s(C, 1)$ and $\{b_i\} = s(C, c_1)$. The read distance of sequence $\{a_i\}$ for $i=2$ is 0, and the read distance of sequence $\{b_i\}$ for $i = 1$ is 0. In addition, $a_2 = b_1 = c_1$, so by Proposition ~\ref{equivsuff} the two sequences are equivalent.
\end{proof}
